``FISTA'' in Banach spaces with adaptive discretisations

### 2022-01-27 response.tex

\newcommand*\justify{%
	\fontdimen2\font=0.4em% interword space
	\fontdimen3\font=0.2em% interword stretch
	\fontdimen4\font=0.1em% interword shrink
	\fontdimen7\font=0.1em% extra space
	\hyphenchar\font=`\-% allowing hyphenation
}
\newcommand{\review}[1]{\texttt{\justify{#1}}}
\newcommand{\todo}[1]{{\color{red}Todo: #1}}
\newcommand{\F}[1]{\mathbb{#1}}
\let\vecf\vec
\renewcommand{\vec}[1]{\mathbf{#1}}
\definecolor{darkgreen}{rgb}{0.0, 0.5, 0.0}
\renewcommand{\ULthickness}{1.0pt}
\newcommand{\edit}[2]{{\ifmmode\text{\color{red}\sout{\ensuremath{#1}}}\else {\color{red} \sout{#1}}\fi} {\color{darkgreen} #2}}

\title{Response to reviewer comments:\\``FISTA'' in Banach spaces with adaptive discretisations}
\author{Antonin Chambolle and Robert Tovey}
\date{7th October 2021}

\begin{document}
\maketitle

We would like to thank both the reviewers again for their time and feedback. A detailed response to each reviewer's comments is given below. All new text is marked in the manuscript and all deletions outside of the appendix. In particular, the statements of Theorem 7 and Lemmas 14-17 have changed so the proofs are now irrelevant. Changes to the theorems are fully annotated, the original proofs are hidden.

We have also made one change which was not in response to the comments of the reviewer. It became clear to us that the statement of Definition 1 is much more natural as a property of the sequence $w_n$ rather than the sets $\F{U}^n$. Morally, we assume the existence of a `nice' minimising sequence rather than asserting that $w_n$ must be minimisers in $\F{U}^n$. Documenting the changes which are longer than one line:
\begin{itemize}
	\item At the end of page 4, emphasis is shifted towards the choice of sequence $w_n$
	\item Definition 1 has been updated
	\item At the top of page 8, the relaxation of Definition 1 allows an extra observation. In particular, the choice $\F{U}^n=\F H$ is now valid in Theorems 2 and 3.
	\item In Theorem 3 we now have the specific choice $\tilde{w}_k = u_{n_k-1}$. The old $a_U=1$ is removed leading to a much simpler proof in the appendix.
	\item The notation of Lemma 6 (and its proof in Lemma 14) has changed without changing the result, namely $(u_k,E(\tilde{w}_k))\mapsto (\tilde{w}_k, E(\tilde{\F{U}}^k))$.
	\item The proof of Theorem 3 (i.e. Theorem 7) is now much simpler.
\end{itemize}

@@

	\item \review{(25) What is $\partial_n E$?}

	Old (25) is now (27). This refers to the derivative of $E$ on the subspace $\F U^n$, this is now clarified immediately above it in the text.

	\item \review{Inequalities (61),(62) do not make sense: the quantities on the left and right-hand sides are equal.}

	Please can the reviewer clarify this point? The old (61-62) are now (63-64) which are equalities. There was a pair of inequalities (59-60), now (61-62), but they are simply statements of H\"older's inequality which we cannot expect to be sharp in general.

### 2022-01-27 paper.tex
General update

	\section{Introduction}
	The Fast Iterative Shrinkage-Thresholding Algorithm (FISTA) was proposed by Beck and Teboulle \cite{Beck2009} as an extension of Nesterov's fast gradient method \cite{Nesterov2004} and is now a very popular algorithm for minimising the sum of two convex functions.
	We write this as the problem of computing
	\begin{equation}\label{eq: min F = fD+fC}
		\Emin \qquad\text{such that}\qquad \op{E}(\var0)\coloneqq \op{f}(\var0) + \op{g}(\var0),
	\end{equation}
	where $\op{f}\colon\F H\to\R$ is a convex differentiable function with $L$-Lipschitz gradient and $\op{g}\colon\F H\to\bar\R$ is a convex function, whose ``proximity operator'' is easy to compute. The iterates of the FISTA algorithm will be denoted $\var0_n\in\F H$. It has been shown that $\op{E}(\var0_n) - \Emin$ converges at the optimum rate of $n^{-2}$ \cite{Beck2009}, and later (after a small modification) the convergence of the iterates was also shown in a general Hilbert space setting \cite{Chambolle2015}. Many further works have gone on to demonstrate faster practical convergence rates for slightly modified variants of FISTA \cite{Tao2016,Liang2017,Alamo2019}.

@@

	\paragraph{Step 2: Quantifying the scaling properties}
	\begin{adaptiveStep}\label{step: step 2 smoothness}
		To show that the extra terms in \eqref{eq: mini FISTA convergence} are small, we need to quantify the approximation properties of $\F{U}^n$. The idea is that $\F{U}^n$ are locally compact, so that minimisers are achieved, but `growing' so that $\min_{\var0\in\F{U}^n}\op{E}(\var0) \to \inf_{\var0\in\F H}\op{E}(\var0)$. A canonical example might be $\F{U}^n=\{\var0\in\F H\st \norm{\var0}\leq n\}$. To quantify the balance between a growing Hilbert norm and improving estimate of $\op{E}$, we introduce a secondary sequence $n_0 < n_1 < \ldots$ and constants $\aU,\aE\geq1$ such that for each $k\in\F N$
		\begin{equation}\label{eq: first regularity}
			n\leq n_k \implies \norm{\var2_n} \lesssim \aU^k, \qquad n\geq n_k \implies \op{E}_0(\var2_n) \lesssim \aE^{-k}.
		\end{equation}
		The expanding balls example would become $\F{U}^n=\{\var0\in\F H\st \norm{\var0}\leq \aU^k\}$ for $n\in[n_k,n_{k+1})$, and $\aE$ reflects the smoothness of $\op{E}$. The choice of exponential scaling \edit{is more}{is introduced to improve stability of Algorithm~\ref{alg: refining FISTA}. It is} natural if we consider the $\F{U}^n$ to be the subspace of functions discretised on a uniform mesh. If that mesh is sequentially refined, then the resolution of the mesh will be of order $h^k$ after $k$ refinements and for some $h<1$. The integer $n_k$ is then the time at which the mesh has refined $k$ times. \edit{The value $\aU$ reflects the smoothness of the chosen basis, and $\aE$ is a measure of efficiency with respect to the structure of $\op{E}$. If an efficient basis is used, then $\aE$ is large.}{} The trade-off between $\aE$ and $\aU$ dictates the final convergence rate of the algorithm. If $\aU>1$, then we cannot guarantee the original $n^{-2}$ rate of convergence.
	\end{adaptiveStep}

	\paragraph{Step 3: Generalising the convergence bound}
	\begin{adaptiveStep}\label{step: step 3 convergence bound}
		In this step we combine the FISTA stability estimate with the subset approximation guarantees to provide a sharper estimate of stability with respect to the parameters $\aE$ and $\aU$. For example, if \edit{$\F{U}^n$ is of the form}{for each $k\in\F N$}
		$$\edit{\F{U}^{n_k} = \F{U}^{n_k+1} = \ldots = \F{U}^{n_{k+1}-1}}{\var2_{n_k}=\var2_{n_k+1}=\ldots \var2_{n_{k+1}-1}},$$
		then many terms on the right-hand side of \eqref{eq: mini FISTA convergence} telescope to 0. The result of this is presented in Lemma~\ref{thm: mini exponential FISTA convergence}. The key idea is that the stability error in \eqref{eq: mini FISTA convergence} has $K\lesssim\log(N)$ terms, rather than $N$.
	\end{adaptiveStep}

@@

	\paragraph{Step 5: Sufficiently slow \edit{refinement}{growth}}
	\begin{adaptiveStep}\label{step: step 5 sufficiently slow}
		The result of Step~\ref{step: step 4 sufficiently fast} is sufficient to prove convergence, but not yet a rate. If the subsets grow too quickly, then the influence of $\norm{\var0_n}\to\infty$ will slow the rate of convergence. \edit{If the resolution is too coarse then we are overfitting to the discretised problem, but if the resolution is too fine then convergence of the energy is slow.}{If $k$ is too small (coarse discretisation), then we overfit to the discrete problem, but if $k$ is too large, then the energy converges slowly.} Lemma~\ref{thm: sufficiently slow} balances these two factors in an optimal way ($n_k^2\simeq \aE^k\aU^{2k}$) for Algorithm~\ref{alg: refining FISTA} resulting in a convergence rate of 
		$$ \op{E}_0(\var0_N) \lesssim \frac{\aU^{2K}}{N^2} \lesssim \frac{N^{2\kappa}}{N^2}$$
		for all $N\in\F N$ and \edit{some }{}$\kappa\edit{}{=\frac{2\log\aU}{\log\aE+2\log\aU}}\in[0,1)$\edit{ depending on $\aU,\aE$}{}. In particular, if the minimum is attained in $\F H$, then we recover the classical rate with $\kappa=0$.
	\end{adaptiveStep}

@@

		Spatial adaptivity is robust, so long as the scaling properties of Step~\ref{step: step 2 smoothness} are satisfied. The only other constraint is to ensure that the partial telescoping of Step~\ref{step: step 3 convergence bound} still holds.%
		}{%
		The main challenge for adaptive choice of $\F{U}^n$ is to guarantee \eqref{eq: first regularity} from Step~\ref{step: step 2 smoothness} using a posteriori estimates. Combined with the partial telescoping requirement in Step~\ref{step: step 3 convergence bound}, a natural choice is $\var2_n=\var0_{n_k-1}$ for $n\in[n_k,n_{k+1}-1)$, i.e. the value of $n_k$ is chosen to be $n+1$ once the iterate $\var0_n$ is observed.
		}
		Theorem~\ref{thm: stronger exponential FISTA convergence} shows that a sufficient condition is
		$$\var0_{n_k-1}\in\F{U}^{n_k}\cap\F{U}^{n_k+1}\cap\ldots\cap\F{U}^{n_{k+1}-1}\edit{.}{, \qquad \norm{\var0_{n_k-1}} \lesssim \aU^k, \quad\text{and}\quad \op{E}_0(\var0_{n_k-1}) \lesssim \aE^{-k}}$$
		Convergence is most stable if the approximation spaces $\F{U}^n$ satisfy a monotone inclusion, breaking the monotonicity requires more care.
		\edit{Adaptively choosing the time of refinement is more challenging due to the non-descent property of FISTA. The idea is that $\F{U}^n$ should be refined as soon as the value $\op{E}(\var0_n)$ is sufficiently close to $\op{E}(\var2_n)$. This is analysed in Theorem~\ref{thm: stronger exponential FISTA convergence} (sufficient conditions in Lemma~\ref{thm: practical refinement criteria}) which shows convergence results of the form}{The only non-trivial property to verify is the energy gap  $\op{E}_0(\var0_n) = \op{E}(\var0_n)-\Emin$. Lemma~\ref{thm: practical refinement criteria}) proposes some sufficient conditions to guarantee the same overall rate of convergence as in Step~\ref{step: step 5 sufficiently slow},}
		$$ \min_{n\leq N}\op{E}_0(\var0_n) \lesssim \frac{N^{2\kappa}}{N^2}$$
		for all $N\in\F N$, with the same $\kappa\in[0,1)$ from Step~\ref{step: step 5 sufficiently slow}. The penalty for accelerating the change of discretisation is a potential loss of stability or monotonicity in $\op{E}(\var0_n)$, although this behaviour has not been seen in numerical experiments.
	\end{adaptiveStep}

@@

	\noindent The proof is given in Lemma~\ref{app:thm: sufficiently fast}. We make two observations of the optimality of Lemma~\ref{thm: sufficiently fast}:
	\begin{itemize}
		\item The convergence guarantee for $N$ iterations of classical FISTA in the space $\F{U}^N$ is 
		$$ \op{E}_0(\var0_N) \lesssim \frac{\norm{\var0_0-\edit{\argmin_{\var0\in\F{U}^N}\op{E}(\var0)}{\tilde{\var2}_K}}^2}{N^2} + \min_{\var0\in\F{U}^N}\op{E}_0(\var0) \lesssim \frac{\aU^{2K}}{N^2} + \aE^{-K}.$$
		This is equivalent to Lemma~\ref{thm: sufficiently fast} after the assumptions on $n_k$.
		\item If $\F H$ is finite dimensional, then the $\aU=1$ is almost trivially satisfied. Norms in finite dimensions are equivalent and any discretisation can be achieved with a finite number of refinements (i.e. the sums over $k$ are finite). 
	\end{itemize}

@@

	\begin{lemma}\label{thm: sufficiently slow}
		\Copy{thm: sufficiently slow}{
			Suppose $\var0_n$ and $n_k$ are sequences satisfying 
			$$\op{E}_0(\var0_N) \lesssim \frac{\aU^{2K}}{N^2} \qquad\text{where} \qquad  n_K^2\gtrsim \aE^K\aU^{2K},$$
			then 
			$$\op{E}_0(\var0_N) \lesssim \frac{1}{N^{2(1-\kappa)}}\qquad \text{ where } \qquad \kappa = \frac{\log \aU^2}{\log \aE + \log \aU^2}.$$}
	\end{lemma}
	\noindent The proof is given in Lemma~\ref{app:thm: sufficiently slow}.

@@

	Theorem~\ref{thm: exponential FISTA convergence} already allows for spatial adaptivity, so we focus on temporal adaptivity. Lemma~\ref{thm: sufficiently fast} suggests that a good refinement time strategy is to choose $n_k$ to be the minimal integer such that $\op{E}_0(\var0_{n_k-1})\lesssim \aE^{-k}$. However, the value of $\op{E}_0$ may be hard to estimate and so we retain a `backstop' condition which guarantees that convergence is no slower than the rate given by Theorem~\ref{thm: exponential FISTA convergence}. \edit{This is confirmed by}{In the non-classical case of $\aU>1$, we provide} the following theorem.
	\begin{theorem}\label{thm: stronger exponential FISTA convergence}
		\Copy{thm: stronger exponential FISTA convergence}{
			Let $(\F{U}^n\subset\F H)_{n\in\F N}$ be a sequence of subsets satisfying \eqref{eq: refining subspace definition}\edit{}{, $(\tilde{\var2}_k)_{k\in\F N}$ an \aUaE,} and $n_k\in\F N$ a monotone increasing sequence such that
			$$ \edit{\tilde{\F{U}}^k \coloneqq \F{U}^{n_k}, \qquad}{} \tilde{\var2}_k \in \edit{\left[\argmin_{\var0\in\tilde{\F{U}}^k}\op{E}(\var0)\right]}{\F{U}^{n_k}}\cap\F{U}^{n_k+1} \cap\ldots\cap \F{U}^{n_{k+1}-1}$$
			for all $k\in\F N$.
			Compute $\var0_n$ and $\var1_n$ by Algorithm~\ref{alg: refining FISTA}.

			\edit[0]{\bigbreak\noindent \sout{Suppose there exist $\aU,\aE\geq1$ such that either:}
			\begin{itemize}
				\item \sout{$\aU>1$ and $n_k^2\lesssim \aE^k\aU^{2k}$, or}
				\item \sout{$\aU=1$, $\sum_{k=1}^\infty n_k^2\aE^{-k}<\infty$ and $\sum_{k=1}^\infty\norm{\tilde{\var2}_k-\tilde{\var2}_{k+1}}<\infty$}
			\end{itemize}
			\sout{and both}
			$$ \norm{\tilde{\var2}_k}\lesssim \aU^k \qquad\text{and}\qquad \op{E}_0(\var0_{n_k-1}) \lesssim \aE^{-k}.$$
			\sout{Whenever these conditions on $n_k$ are satisfied, }}{If $\aU>1$, $\aE\geq1$, and $n_k^2\lesssim \aE^k\aU^{2k}$, }then
			$$\min_{n\leq N}\op{E}_0(\var0_n) \lesssim \frac{1}{N^{2(1-\kappa)}}\qquad \text{ where } \qquad \kappa = \frac{\log \aU^2}{\log \aE + \log \aU^2}$$
			uniformly for $N\in\F N$.
		}
	\end{theorem}
	\noindent The proof is given in Theorem~\ref{app:thm: stronger exponential FISTA convergence}. If we directly compare Theorems~\ref{thm: exponential FISTA convergence} and~\ref{thm: stronger exponential FISTA convergence}, we note that the convergence rate is the same in both theorems but the price for better adaptivity (i.e. only an upper bound on $n_k$) is a slightly weaker stability guarantee (now convergence of $\min_{n\leq N}\op{E}_0(\var0_n)$). In Theorem~\ref{thm: exponential FISTA convergence}, as in the original FISTA algorithm, the sequence $\op{E}(\var0_n)$ is not monotone but the magnitude of oscillation is guaranteed to decay in time. This behaviour is lost in Theorem~\ref{thm: stronger exponential FISTA convergence}. Although we do not prove it here, it can be shown that the stronger condition 
	\begin{equation}
		\tilde{\var2}_k \in \edit{\left[\argmin_{\var0\in\tilde{\F{U}}^k}\op{E}(\var0)\right]}{\F{U}^{n_k}}\cap\F{U}^{n_k+1} \cap\ldots\cap \F{U}^{n_{k+1}-1}\cap\ldots\cap \F{U}^N
	\end{equation}
	is sufficient to restore the stronger last-iterate guarantee on $\op{E}_0(\var0_N)$. Again, monotonicity of $\F{U}^n$ corresponds with improved stability of Algorithm~\ref{alg: refining FISTA}.

>>>>

	\section{Introduction}
	The Fast Iterative Shrinkage-Thresholding Algorithm (FISTA) was proposed by Beck and Teboulle \cite{Beck2009} as an extension of Nesterov's fast gradient method \cite{Nesterov2004} and is now a very popular algorithm for minimising the sum of two convex functions.
	We write this as the problem of computing
	\begin{equation}\label{eq: min F = fD+fC}
		\Emin \qquad\text{such that}\qquad \op{E}(\var0)\coloneqq \op{f}(\var0) + \op{g}(\var0),
	\end{equation}
	where $\op{f}\colon\F H\to\R$ is a convex differentiable function with $L$-Lipschitz gradient and $\op{g}\colon\F H\to\bar\R$ is a convex function, whose ``proximity operator'' is easy to compute. \edit{}{Throughout this work we assume that $\op{E}$ is bounded below so that the infimum is finite. }The iterates of the FISTA algorithm will be denoted $\var0_n\in\F H$. It has been shown that $\op{E}(\var0_n) - \Emin$ converges at the optimum rate of $n^{-2}$ \cite{Beck2009}, and later (after a small modification) the convergence of the iterates was also shown in a general Hilbert space setting \cite{Chambolle2015}. Many further works have gone on to demonstrate faster practical convergence rates for slightly modified variants of FISTA \cite{Tao2016,Liang2017,Alamo2019}.

@@

	\paragraph{Step 2: Quantifying the scaling properties}
	\begin{adaptiveStep}\label{step: step 2 smoothness}
		To show that the extra terms in \eqref{eq: mini FISTA convergence} are small, we need to quantify the approximation properties of $\F{U}^n$. \edit{The idea is that $\F{U}^n$ are locally compact, so that minimisers are achieved, but `growing' so that $\min_{\var0\in\F{U}^n}\op{E}(\var0) \to \inf_{\var0\in\F H}\op{E}(\var0)$.  A canonical example might be $\F{U}^n=\{\var0\in\F H\st \norm{\var0}\leq n\}$.}{The idea is that there is a sequence $\var2_n\in\F{U}^n$, $n\in\F N$ such that $\norm{\var2_n}$ grows slowly and $\op{E}_0(\var2_n)$ decreases quickly.} To quantify \edit{the}{this} balance\edit{ between a growing Hilbert norm and improving estimate of $\op{E}$}{}, we introduce a secondary sequence $n_0 < n_1 < \ldots$ and constants $\aU,\aE\geq1$ such that for each $k\in\F N$
		\begin{equation}\label{eq: first regularity}
			n\leq n_k \implies \norm{\var2_n} \lesssim \aU^k, \qquad n\geq n_k \implies \op{E}_0(\var2_n) \lesssim \aE^{-k}.
		\end{equation}
		\edit{The expanding balls example would become}{A canonical example would be} $\F{U}^n=\{\var0\in\F H\st \norm{\var0}\leq \aU^k\}$ for $n\in[n_k,n_{k+1})$, then $\aE$ reflects the smoothness of $\op{E}$. The choice of exponential scaling \edit{is more}{is introduced to improve stability of Algorithm~\ref{alg: refining FISTA}. It is} natural if we consider the $\F{U}^n$ to be the subspace of functions discretised on a uniform mesh. If that mesh is sequentially refined, then the resolution of the mesh will be of order $h^k$ after $k$ refinements and for some $h<1$. The integer $n_k$ is then the time at which the mesh has refined $k$ times. \edit{The value $\aU$ reflects the smoothness of the chosen basis, and $\aE$ is a measure of efficiency with respect to the structure of $\op{E}$. If an efficient basis is used, then $\aE$ is large.}{} The trade-off between $\aE$ and $\aU$ dictates the final convergence rate of the algorithm. If $\aU>1$, then we cannot guarantee the original $n^{-2}$ rate of convergence.
	\end{adaptiveStep}

	\paragraph{Step 3: Generalising the convergence bound}
	\begin{adaptiveStep}\label{step: step 3 convergence bound}
		In this step we combine the FISTA stability estimate with the subset approximation guarantees to provide a sharper estimate of stability with respect to the parameters $\aE$ and $\aU$. For example, if \edit{$\F{U}^n$ is of the form}{for each $k\in\F N$}
		$$\edit{\F{U}^{n_k} = \F{U}^{n_k+1} = \ldots = \F{U}^{n_{k+1}-1}}{\var2_{n}= \var2_{n_k} \text{ for each } n=n_k,n_k+1,\ldots,n_{k+1}-1},$$
		then many terms on the right-hand side of \eqref{eq: mini FISTA convergence} telescope to 0. The result of this is presented in Lemma~\ref{thm: mini exponential FISTA convergence}. The key idea is that the stability error in \eqref{eq: mini FISTA convergence} has $K\edit{\lesssim\log}{\ll} N$ terms, rather than $N$.
	\end{adaptiveStep}

@@

	\paragraph{Step 5: Sufficiently slow \edit{refinement}{growth}}
	\begin{adaptiveStep}\label{step: step 5 sufficiently slow}
		The result of Step~\ref{step: step 4 sufficiently fast} is sufficient to prove convergence, but not yet a rate. If the subsets grow too quickly, then the influence of $\norm{\var0_n}\to\infty$ will slow the rate of convergence. \edit{If the resolution is too coarse then we are overfitting to the discretised problem, but if the resolution is too fine then convergence of the energy is slow.}{If $n_k$ is too large, then we overfit to the discrete problem, but if $n_k$ is too small, then FISTA converges slowly.} Lemma~\ref{thm: sufficiently slow} balances these two factors in an optimal way ($n_k^2\simeq \aE^k\aU^{2k}$) for Algorithm~\ref{alg: refining FISTA} resulting in a convergence rate of 
		$$ \op{E}_0(\var0_N) \lesssim \frac{\aU^{2K}}{N^2} \lesssim \frac{N^{2\kappa}}{N^2}$$
		for all $N\in\F N$ and \edit{some }{}$\kappa\edit{}{=\frac{2\log\aU}{\log\aE+2\log\aU}}\in[0,1)$\edit{ depending on $\aU,\aE$}{}. In particular, if the minimum is attained in $\F H$, then we recover the classical rate with $\kappa=0$.
	\end{adaptiveStep}

@@

		Spatial adaptivity is robust, so long as the scaling properties of Step~\ref{step: step 2 smoothness} are satisfied. The only other constraint is to ensure that the partial telescoping of Step~\ref{step: step 3 convergence bound} still holds.%
		}{%
		The main challenge for adaptive choice of $\F{U}^n$ is to guarantee \eqref{eq: first regularity} from Step~\ref{step: step 2 smoothness} using a posteriori estimates. Combined with the partial telescoping requirement in Step~\ref{step: step 3 convergence bound}, a natural choice is $\var2_n=\var0_{n_k-1}$ for $n\in[n_k,n_{k+1})$, i.e. the value of $n_k$ is chosen to be $n+1$ once the iterate $\var0_n$ is observed.
		}
		Theorem~\ref{thm: stronger exponential FISTA convergence} shows that a sufficient condition is
		$$\var0_{n_k-1}\in\F{U}^{n_k}\cap\F{U}^{n_k+1}\cap\ldots\cap\F{U}^{n_{k+1}-1}\edit{.}{, \qquad \norm{\var0_{n_k-1}} \lesssim \aU^k, \quad\text{and}\quad \op{E}_0(\var0_{n_k-1}) \lesssim \aE^{-k}}$$
		Convergence is most stable if the approximation spaces $\F{U}^n$ satisfy a monotone inclusion, breaking the monotonicity requires more care.
		\edit{Adaptively choosing the time of refinement is more challenging due to the non-descent property of FISTA. The idea is that $\F{U}^n$ should be refined as soon as the value $\op{E}(\var0_n)$ is sufficiently close to $\op{E}(\var2_n)$. This is analysed in Theorem~\ref{thm: stronger exponential FISTA convergence} (sufficient conditions in Lemma~\ref{thm: practical refinement criteria}) which shows convergence results of the form}{The only non-trivial property to verify is the energy gap  $\op{E}_0(\var0_n) = \op{E}(\var0_n)-\Emin$. Lemma~\ref{thm: practical refinement criteria} proposes some sufficient conditions to guarantee the same overall rate of convergence as in Step~\ref{step: step 5 sufficiently slow},}
		$$ \min_{n\leq N}\op{E}_0(\var0_n) \lesssim \frac{N^{2\kappa}}{N^2}$$
		for all $N\in\F N$, with the same $\kappa\in[0,1)$ from Step~\ref{step: step 5 sufficiently slow}. The penalty for accelerating the change of discretisation is a potential loss of stability or monotonicity in $\op{E}(\var0_n)$, although this behaviour has not been seen in numerical experiments.
	\end{adaptiveStep}

@@

	\noindent The proof is given in Lemma~\ref{app:thm: sufficiently fast}. We make two observations of the optimality of Lemma~\ref{thm: sufficiently fast}:
	\begin{itemize}
		\item The convergence guarantee for $N$ iterations of classical FISTA in the space $\F{U}^N$ is 
		$$ \op{E}_0(\var0_N) \lesssim \frac{\norm{\var0_0-\edit{\argmin_{\var0\in\F{U}^N}\op{E}(\var0)}{\tilde{\var2}_K}}^2}{N^2} + \min_{\var0\in\F{U}^N}\op{E}_0(\var0) \lesssim \frac{\aU^{2K}}{N^2} + \aE^{-K}.$$
		This is equivalent to Lemma~\ref{thm: sufficiently fast} after the assumptions on $n_k$.
		\item If $\F H$ is finite dimensional, then the condition $\aU=1$ is almost trivially satisfied. Norms in finite dimensions are equivalent and any discretisation can be achieved with a finite number of refinements (i.e. the sums over $k$ are finite). 
	\end{itemize}

@@

	\begin{lemma}\label{thm: sufficiently slow}
		\Copy{thm: sufficiently slow}{
			Suppose $\var0_n$ and $n_k$ are sequences satisfying 
			$$\forall n_K\leq N< n_{K+1},\ \op{E}_0(\var0_N) \lesssim \frac{\aU^{2K}}{N^2} \qquad\text{where} \qquad  n_K^2\gtrsim \aE^K\aU^{2K},$$
			then 
			$$\op{E}_0(\var0_N) \lesssim \frac{1}{N^{2(1-\kappa)}}\qquad \text{ where } \qquad \kappa = \frac{\log \aU^2}{\log \aE + \log \aU^2}.$$}
	\end{lemma}
	\noindent The proof is given in Lemma~\ref{app:thm: sufficiently slow}.

@@

	Theorem~\ref{thm: exponential FISTA convergence} already allows for spatial adaptivity, so we focus on temporal adaptivity. Lemma~\ref{thm: sufficiently fast} suggests that a good refinement time strategy is to choose $n_k$ to be the minimal integer such that $\op{E}_0(\var0_{n_k-1})\lesssim \aE^{-k}$. However, the value of $\op{E}_0$ may be hard to estimate and so we retain a `backstop' condition which guarantees that convergence is no slower than the rate given by Theorem~\ref{thm: exponential FISTA convergence}. \edit{This is confirmed by}{In the non-classical case of $\aU>1$, we provide} the following theorem.
	\begin{theorem}\label{thm: stronger exponential FISTA convergence}
		\Copy{thm: stronger exponential FISTA convergence}{
			Let $(\F{U}^n\subset\F H)_{n\in\F N}$ be a sequence of subsets satisfying \eqref{eq: refining subspace definition} \edit{and}{, compute $\var0_n$ and $\var1_n$ by Algorithm~\ref{alg: refining FISTA}. Suppose that there exists a monotone increasing sequence } $n_k\in\F N$ \edit{a monotone increasing sequence }{}such that
			$$ \edit{\tilde{\F{U}}^k \coloneqq \F{U}^{n_k}, \qquad \tilde{\var2}_k \in \left[\argmin_{\var0\in\tilde{\F{U}}^k}\op{E}(\var0)\right]}{\tilde{\var2}_k\coloneqq \var0_{n_k-1}\in\F{U}^{n_k}}\cap\F{U}^{n_k+1} \cap\ldots\cap \F{U}^{n_{k+1}-1}$$
			for all $k\in\F N$.
			\edit{Compute $\var0_n$ and $\var1_n$ by Algorithm~\ref{alg: refining FISTA}.}{}

			\edit[0]{\bigbreak\noindent \sout{Suppose there exist $\aU,\aE\geq1$ such that either:}
			\begin{itemize}
				\item \sout{$\aU>1$ and $n_k^2\lesssim \aE^k\aU^{2k}$, or}
				\item \sout{$\aU=1$, $\sum_{k=1}^\infty n_k^2\aE^{-k}<\infty$ and $\sum_{k=1}^\infty\norm{\tilde{\var2}_k-\tilde{\var2}_{k+1}}<\infty$}
			\end{itemize}
			\sout{and both}
			$$ \norm{\tilde{\var2}_k}\lesssim \aU^k \qquad\text{and}\qquad \op{E}_0(\var0_{n_k-1}) \lesssim \aE^{-k}.$$
			\sout{Whenever these conditions on $n_k$ are satisfied, }}{If $(\tilde{\var2}_k)_{k\in\F N}$ is an \aUaE\ with $\aU>1$ and $n_k^2\lesssim \aE^k\aU^{2k}$, }then
			$$\min_{n\leq N}\op{E}_0(\var0_n) \lesssim \frac{1}{N^{2(1-\kappa)}}\qquad \text{ where } \qquad \kappa = \frac{\log \aU^2}{\log \aE + \log \aU^2}$$
			uniformly for $N\in\F N$.
		}
	\end{theorem}
	\noindent The proof is given in Theorem~\ref{app:thm: stronger exponential FISTA convergence}. If we directly compare Theorems~\ref{thm: exponential FISTA convergence} and~\ref{thm: stronger exponential FISTA convergence}, we note that the convergence rate is the same in both theorems but the price for better adaptivity (i.e. only an upper bound on $n_k$) is a slightly weaker stability guarantee (now convergence of $\min_{n\leq N}\op{E}_0(\var0_n)$). In Theorem~\ref{thm: exponential FISTA convergence}, as in the original FISTA algorithm, the sequence $\op{E}(\var0_n)$ is not monotone but the magnitude of oscillation is guaranteed to decay in time. This behaviour is lost in Theorem~\ref{thm: stronger exponential FISTA convergence}. Although we do not prove it here, it can be shown that the stronger condition 
	\begin{equation}
		\tilde{\var2}_k \in \edit{\left[\argmin_{\var0\in\tilde{\F{U}}^k}\op{E}(\var0)\right]}{\F{U}^{n_k}}\cap\F{U}^{n_k+1} \cap\ldots\cap \F{U}^{n_{k+1}-1}\cap\ldots\cap \F{U}^N
	\end{equation}
	is sufficient to restore the stronger last-iterate guarantee on $\op{E}_0(\var0_N)$. Again, monotonicity of $\F{U}^n$ corresponds with improved stability of Algorithm~\ref{alg: refining FISTA}.

	To enable a more practical implementation of Theorem~\ref{thm: stronger exponential FISTA convergence}, the following lemma describes several refinement time strategies which provide sufficient condition for $\op{E}_0(\edit{\var0_n}{\tilde{\var2}_k}) \lesssim \aE^{-k}$.
	\begin{lemma}\label{thm: practical refinement criteria}
		\Copy{thm: practical refinement criteria}{
			Let \edit{$(\tilde{\F{U}}^k)_{k\in\F N}$ be a sequence of subsets with some points $\var0_k\in \tilde{\F{U}}^k$ and $\tilde{\var2}_k\in\argmin_{\var0\in\tilde{\F{U}}^k}\op{E}(\var0)$. Suppose that }{$(\tilde{\var2}_k)_{k\in\F N}$ be a sequence in $\F H$ with} $\norm{\tilde{\var2}_k}\lesssim \aU^k$. \edit{}{Suppose that $\tilde{\var2}_k\in\tilde{\F U}^k\coloneqq\F{U}^{n_k}$ and denote $\op{E}_0(\tilde{\F{U}}^k) \coloneqq \inf_{\var0\in\tilde{\F U}^k}\op{E}_0(\var0)$. }Any of the following conditions are sufficient to show that $\edit{\{\tilde{\F{U}}^k\}}{\tilde{\var2}_k}$ is an \edit{$(\aU,\aE)$-discretisation for $\op{E}$}{\aUaE}:
			\begin{enumerate}
				\item Small continuous gap refinement: $ \op{E}_0(\edit{\var0_k}{\tilde{\var2}_k})\leq \vars1\aE^{-k}$ for all $k\in\F N$, some $\vars1>0$.
				\item Small discrete gap refinement: $\op{E}_0(\edit{\tilde{\var2}_k}{\tilde{\F{U}}^k})\leq \vars1\aE^{-k}$ and $ \op{E}_0(\edit{\var0_k}{\tilde{\var2}_k})-\op{E}_0(\edit{\tilde{\var2}_{k-1}}{\tilde{\F{U}}^{k-1}})\leq \vars1\aE^{-k}$ for all $k\in\F N$, some $\vars1>0$.
				\edit[2]{\item Small relative gap refinement: $ \op{E}_0(\edit{\var0_k}{\tilde{\var2}_k}) - \op{E}_0(\edit{\tilde{\var2}_{k-1}}{\tilde{\F{U}}^{k-1}}) \leq \vars1\op{E}_0(\edit{\var0_k}{\tilde{\var2}_k})$ for all $k\in\F N$, some $0<\vars1\leq \frac{1}{1+\aE}$.}{}
			\end{enumerate}
			\edit{If}{Otherwise, suppose} there exists a Banach space $(\F{U},\Norm\cdot)$ \edit{with $\F{U}\supset\F H$}{which contains each $\tilde{\F{U}}^k$, $\sup_{k\in\F N}\Norm{\tilde{\var2}_k}<\infty$,} and the sublevel sets of $\op{E}$ are $\Norm\cdot$-bounded, then it is also sufficient that:
			\begin{enumerate}\setcounter{enumi}{2}
				\item Small continuous gradient refinement: $ \inf_{\var1\in\partial\op{E}(\edit{\var0_k}{\tilde{\var2}_k})}\edit{\Norm{\var1}_*}{\sup_{\var0\in\F{U}} \frac{|\IP{\var1}{\var0}|}{\Norm{\var0}}}\leq \vars1\aE^{-k}$ for all $k\in\F N$, some $\vars1>0$.
				\item Small discrete gradient refinement: $\op{E}_0(\edit{\tilde{\var2}_k}{\tilde{\F{U}}^k})\leq \vars1\aE^{-k}$ and $\sup_{\var0,\tilde{\var2}\in \tilde{\F{U}}^k}\inf_{\var1\in\partial\op{E}(\edit{\var0_k}{\tilde{\var2}_k})}\frac{|\IP{v}{\var0-\tilde{\var2}}|}{\Norm{\var0-\tilde{\var2}}}\leq \vars1\aE^{-k}$ for all $k\in\F N$, some $\vars1>0$.
			\end{enumerate}
		}
	\end{lemma}
